---
title: "Declan's Fitbit Data - July 2021"
author: "Declan Murphy"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

```{r setup, include = FALSE}
rm(list=ls())
setwd("/Users/declanmurphy/Documents/Health")
library(tidyverse)
library(readxl)
library(lubridate)
```

## Data Import
I downloaded this past month's fitness data, measured by my Fitbit Inspire 2,
to take a look at some trends and areas for potential improvement. These records
are provided in a handy Excel format for the past 31 days.

Note that I purchased by Fitbit on 7/3/21, so I'm missing data for 7/1 and 7/2.

```{r Data Import}
Fitbit <- read_excel("/Users/declanmurphy/Documents/Health/fitbit_export_20210723..xls",
                     sheet = 3)

Fitbit$Date <- ymd(Fitbit$Date)
```

We've brought in the data set, specifying sheet 3 to capture only the data on 
my activity. Now let's take a look at a preview of the data:

```{r Preview}
print(Fitbit)
```
## Additional Variables for Analysis
The activity data is imported! But before we do any analysis, we need to add some new
variables -- namely, a variable for TOTAL minutes that are active, and a variable
for calories burned due to my base metabolic rate (BMR). So let's create these easily:

```{r New Variables}
Fitbit <- Fitbit %>%
  mutate(
    Total_Minutes_Active = `Minutes Lightly Active` + `Minutes Fairly Active` +
      `Minutes Very Active`,
    BMR_Calories = `Calories Burned` - `Activity Calories`)

print(Fitbit)
```

## Graph 1: Active Minutes
First, let's see how my active minutes varied over the course of the month.
```{r Graphing Active Minutes}
Active_graph <- ggplot(data=Fitbit,
                       aes(x=`Date`, y=`Total_Minutes_Active`))+
  geom_bar(stat="identity",
           fill = "dodgerblue") +
  labs(title = "Active Minutes for July",
       y = "Total Active Minutes")+
  theme_minimal()

print(Active_graph)

```

## Graph 2: Calories Burned
```{r Graphing Calories Burned}
Calories <- ggplot(Fitbit,
                   aes(x= `Date`, y = `Calories Burned`)) +
  geom_line(color = "red") +
  geom_point(shape = 17) +
  geom_hline(yintercept = 3700, 
             color = "blue") +
  scale_x_date(date_breaks = "5 days") +
  theme_minimal()

print(Calories)

```


## Graph 3: Burn Efficiency
Since we've already looked at active minutes and calories burned, the natural 
next question is: how well did these minutes of activity translate into calories?

For this first bar graph, we can use the existing variables to generate a new 
vector, `Burn Rate`, which is simply the average calories burned per minute of 
activity.

```{r Simple Bar}

Fitbit <- Fitbit %>%
  mutate(Burn_Rate = `Activity Calories` / `Total_Minutes_Active`)

Avg_Burn <- ggplot(Fitbit, 
                   aes(x=`Date`, y=Burn_Rate))+
  geom_bar(stat="identity",
           fill = "dodgerblue")+
  labs(y = "Calories Burned Per Minute of Activity")+
  theme_minimal()
print(Avg_Burn)
```

## Graph 4: Linear Fit of Calories Burned vs. Active Minutes

```{r Calories vs. Minutes}
Burn2 <- ggplot(Fitbit, aes(x=`Total_Minutes_Active`,
                           y=`Activity Calories`))+
  geom_smooth(method = "lm")+
  geom_point()+
  labs(x = "Total Active Minutes")+
  theme_minimal()

print(Burn2)
```

